---
title: "Problema 2.5"
author: "Mau"
date: "2024-02-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problema 2.5:

Original Text:

Consider a random walk on {0,...,k} ,which moves left and right with respective probabilities q and p. If the walk is at 0 it transitions to 1 on the next step. If the walk is at k it transitions to k - 1 on the next step. This is called random walk with relecting boundaries. Assume that k = 3, q = 1∕4, p = 3∕4, and the initial distribution is uniform. For the following, use technology if needed.

Traducido:

Considere una caminata aleatoria en {0,...,k} ,que se mueve hacia la izquierda y hacia la derecha con las respectivas probabilidades q y p. Si la caminata está en 0, pasa a 1 en el siguiente paso. Si la caminata está en k, pasa a k - 1 en el siguiente paso. A esto se le llama paseo aleatorio con límites de reelección. Supongamos que k = 3, q = 1∕4, p = 3∕4 y que la distribución inicial es uniforme. Para lo siguiente, use la tecnología si es necesario.

(a) Exhibit the transition matrix.

```{r}
P<-matrix(c(0,1,0,0,
         1/4,0,3/4,0,
         0,1/4,0,3/4,
         0,0,1,0),nrow = 4,byrow = TRUE)
P
```

(b) Find $P(X_7 = 1|X_0 = 3,X_2 = 2,X_4 = 2)= P(X_7 = 1|X_4 = 2)$

P\^3
